---
title: "Section Worksheet: Regular Expressions"
author: 
date: "March 19, 2015"
output: pdf_document
---

$ ends the match
[[:alpha:]]"@[[:alpha:]]$
matches c@t but not c@ts

[[:alpha:]]+ matches letter 1 or more times
[[:alpha:]]* matches letter 0 or more times




Write down a regular expression to match the following:

* Words with @ symbols in them, e.g., vi@gra

* An IP address (4 sets of 1-3 digits separated by periods, e.g., 124.32.6.240)
* A typical email address that ends with .com, .edu, .net, .org, or .gov
grep("^[0-9]{1,3}\\.[0-9]{1,3}\\.[0-9]{1,3}\\.[0-9]{1,3}",ip addresses)


alnum=[0-9][a-z][A-Z]
grep(ˆ[[:alnum:]_\\.]+@[[:alnum:]]+\\(edu|com|gov|net|com\\)$, movies)

Consider the following character vector,
```{r, echo = FALSE}
movies <- c("The Shawshank Redemption (1994)"   ,
            "The Godfather (1972)" ,
            "The Godfather: Part II (1974)",
            "Pulp Fiction (1994)",
            "The Good, the Bad and the Ugly (1966)",
            "12 Angry Men (1957)")
```
```{r} 
movies
```
What is the return expression from each of the following function calls:
```{r, eval = FALSE}
grep("I{2,}", movies)
```
The goodfather movie. it will match anything that has 2 I's. the brackets work to indicate {min, max} in terms of number of repetitions.
```{r, eval = FALSE}
grep("Go+d", movies)
```
will retrieve anythingt that has Go at the bgeining and d at the end, not mattering the content or length of what is in between. 
```{r, eval = FALSE}
gregexpr("\\(.*\\)", movies[1])
```
\\ gets rid of the special meaning
\\( sees "(" as a string rather then the special meaning
.- matches single character (it is a wild character, matching ANY character)
*- we repeat the character class to the left 0 or more times

gregexpr es una function de location it tells you where the first parenthesis is


```{r, eval = FALSE}
gsub("[0-9]", "", movies[6])
```
will match any number of length 1. by using gsub will let us replace strings with something else. in this case you replace all numbers with a blank. 
in gsub("[0-9]+", "", "movies12")= moviesp
gsugsub("[0-9]", "", "movies12")= moviespp
```{r, eval = FALSE}
gsub("[[:blank:]].*", "", movies[5])
```
$ means that the string needs to end inmediately after the preceding element. 
blank will match a white space. 
so this matches a space followed by anything. 
so it deletes everything after the first space. 

```{r, eval = FALSE}
gsub(" \\(.*", "", movies[5])
```


 Suppose we want to match the word 'cat' or 'at' or 't' but
 don't want to match 'cat' embedded within another word
 There can be other words present:
```{r}
cats = c("diplocat", "Hi cat", "mat", "at", "t!", "ct")
```
The \< stands for beginning of a word and
 \> stands for the end of a word
In R we have to escape the \\ with an extra \\.
```{r}
grep("\\<(cat|at|t)\\>", cats)
grep("\\<(ca|a)?t\\>", cats)
```

The following do not work as expected can you figure out why?
```{r}
grep("\\<c?a?t)\\>", cats)
grep("^(cat|at|t)$", cats)
```
? always means that the preceding character is optional to match. 
ojo= (oa)? would take matching oa as optional, but [oa]? would make o and a elements of a set and would take EITHER LETTER as optional. 


Find the word cat or caat or caaat, etc.
```{r}
caats = c("cat", "caat.", "caats", "caaaat", "my cat")
grep("\\<+t\\>", caats)
# the {1,} is equivalent to +
grep("\\<ca{1,}t\\>", caats)
```

 Now we want to find dog anywhere in the string
 We don't care about capitals
```{r}
dogs = c("dogmatic", "TopDog","Doggone it!", "RUN DOG RUN")
# The tolower function is handy here.
grep("dog", tolower(dogs))
grep("[Dd][Oo][Gg]", dogs)
```

Finally we are looking at character vectors where
 each entry must be a number.
 The number can have an optional sign in front of it
 The number can have an optional decimal point followed by digits
```{r}
nums = c("1.2", "-3000", "5lo", "hi2", "12.", "+57")
grep("^[-+]?[[:digit:]]+(\\.[[:digit:]]+)?$", nums)
```
ˆ means that you have to start matching stuff from the begining of the string and not in the middle. 
[-+]? accounts for an optional plus or minus sign at the beggining of the expression. 
[[:digit:]]+ matches one or more digits (study the + sign!)
\\ gets rid of the special meaning of the symbol used in regular expressions (put it BEFORE the symbol that you want to have LITERALLY)... in this case is a dot (to represent decimal places)
. means that you can match any element
[[:digit:]]+ again, means matches one or more digits
? makes all the ultimate part optional 
